---
title: "Plot TrackMate Features"
output: html_notebook
---
Add libraries and generate data
```{r}
#this script reads mean and std error values from trackmate data
#run through a matlab script (morphChange2) and plots various graphs

library(ggplot2);
library(dplyr);
library(plotly);
library(ggpubr);
library(cowplot);
library(plotrix);
library(data.table);
library(FNN);
library(effsize);

# select first .csv feature file to examine
fname <- file.choose()
wd <- getwd()
setwd(wd)

featureData <- read.csv(fname)
# print names of all features in the csv
colnames(featureData)
# determine which feature to graph (hard-coded, change the col value)
selectedCol = 29

# read units (if they exist)
# if col is DirectedMigration (has no units), take from displacement units
# broken bc cant read mu
# if(selectedCol == 29){
#   units = featureData[3,8]
#   units <- substr(units, 2, nchar(units)-1)
# } else if(df[4, selectedCol]!=''){
#   units = featureData[3,selectedCol]
#   units <- substr(units, 2, nchar(units)-1)
# } else{
#   units = ''
# }

# read units from 4th row
selectedFeatureData <- subset(featureData, select = selectedCol)
# ADD FILTERS IF NECESSARY
selectedFeatureData <- selectedFeatureData[-c(1:3),, drop=FALSE]

# select second .csv feature file to examine
fname <- file.choose()

featureData2 <- read.csv(fname)

# read units from 4th row
selectedFeatureData2 <- subset(featureData2, select = selectedCol)
selectedFeatureData2 <- selectedFeatureData2[-c(1:3),, drop=FALSE]

# remove NA and negative values
selectedFeatureData <- selectedFeatureData[complete.cases(selectedFeatureData),]
selectedFeatureData <- read.table(text=selectedFeatureData,col.names = c("Scratch"))
selectedFeatureData <- selectedFeatureData[selectedFeatureData$Scratch>0,]
selectedFeatureData <- read.table(text=as.character(selectedFeatureData),col.names = c("Scratch"))
selectedFeatureData2 <- selectedFeatureData2[complete.cases(selectedFeatureData2),]
selectedFeatureData2 <- read.table(text=selectedFeatureData2,col.names = c("PDMS"))
selectedFeatureData2 <- selectedFeatureData2[selectedFeatureData2$PDMS>0,]
selectedFeatureData2 <- read.table(text=as.character(selectedFeatureData2),col.names = c("PDMS"))
```
Remove outliers
```{r}
# Remove outliers
mean_scr <- mean(selectedFeatureData$Scratch)
sd_scr <- sd(selectedFeatureData$Scratch)

lbound <- mean_scr - 2* sd_scr
ubound <- mean_scr + 2* sd_scr

filteredScrDM <- selectedFeatureData[selectedFeatureData$Scratch >= lbound & selectedFeatureData$Scratch <= ubound,]

write.csv(filteredScrDM, paste(wd, "filtered_scratch_DM.csv", sep = "/"), row.names=FALSE)

mean_pdms <- mean(selectedFeatureData2$PDMS)
sd_pdms <- sd(selectedFeatureData2$PDMS)

lbound <- mean_pdms - 2* sd_pdms
ubound <- mean_pdms + 2* sd_pdms

filteredPDMSDM <- selectedFeatureData2[selectedFeatureData2$PDMS >= lbound & selectedFeatureData2$PDMS <= ubound,]

write.csv(filteredPDMSDM, paste(wd, "filtered_pdms_DM.csv", sep = "/"), row.names=FALSE)

# Calculate effect size
cohen.d(filteredScrDM, filteredPDMSDM)
```

Generate bar plot with error bars

```{r}
# convert character dataframe type to numeric, hard-coded, fix later, issue converting string to column name
selectedFeatureData <- transform(selectedFeatureData, DirectedMigration=as.numeric(DirectedMigration))
selectedFeatureData2 <- transform(selectedFeatureData2, DirectedMigration=as.numeric(DirectedMigration))

# Sample data
materials <- c("Scratch", "PDMS")
means <- c(as.numeric(colMeans(selectedFeatureData, na.rm=TRUE)), as.numeric(colMeans(selectedFeatureData2, na.rm=TRUE))) 
stdes <- c(as.numeric(std.error(selectedFeatureData, na.rm=TRUE)), as.numeric(std.error(selectedFeatureData2, na.rm=TRUE))) 

# Create dataframe
featureStats <- data.frame(
  Material = materials,
  Mean = means,
  StdError = stdes
)

# Create the plot
ggplot(featureStats, aes(x = Material, y = Mean, fill = Material)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_errorbar(aes(ymin = Mean - StdError, ymax = Mean + StdError), 
                width = 0.2, position = position_dodge(0.7)) +
  labs(x = "Material",
       y = "Directed Migration") +
  #sets theme of plot
  theme(
    legend.position="none",
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color="black", fill="NA", size=1),
    panel.grid.major = element_line(size = 1, linetype = 'solid', color = "white"),
    panel.grid.minor = element_line(size = 0.5, linetype = 'solid', color = "white"),
    # LABLES APPEARANCE
    plot.title = element_text(size=20, face= "bold", colour= "black"),
    axis.title.x = element_text(size=12, face="bold", colour = "black"),    
    axis.title.y = element_text(size=12, face="bold", colour = "black"),    
    axis.text.x = element_text(size=10, face="bold", colour = "black"),
    axis.text.y = element_text(size=10, face="bold", colour = "black"),
    legend.text = element_text(size=12, face="bold", colour = "black"),
    legend.title  = element_text(size=12, face="bold", colour = "black"),
  ) 
```
# plot bar graph of two selected datasets
plot <- ggplot(comMean, aes(x=colNames(featureStats), y=DirectedMigration)) +
  #plots bar graph based on number in dataframe vs counting an amt of some variable
  geom_bar(aes(x=Treatment_Group, y=Velocity,
                  fill=Treatment_Group),
               stat="identity") +
  #plots error based on calculated values
  geom_errorbar(aes(x=colNames(featureStats), 
                    ymin=DirectedMigration-, 
                    ymax=DirectedMigration+std$V1),
                colour="black",
                width=0.4) +
  #sets theme of plot
  theme(
    legend.position="none",
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color="black", fill="NA", size=1),
    panel.grid.major = element_line(size = 1, linetype = 'solid', color = "white"),
    panel.grid.minor = element_line(size = 0.5, linetype = 'solid', color = "white"),
    # LABLES APPEARANCE
    plot.title = element_text(size=20, face= "bold", colour= "black"),
    axis.title.x = element_text(size=12, face="bold", colour = "black"),    
    axis.title.y = element_text(size=12, face="bold", colour = "black"),    
    axis.text.x = element_text(size=10, face="bold", colour = "black"),
    axis.text.y = element_text(size=10, face="bold", colour = "black"),
    legend.text = element_text(size=12, face="bold", colour = "black"),
    legend.title  = element_text(size=12, face="bold", colour = "black"),
  ) +
  #sets labels. hard-coded
  labs (x="Treatment Group", 
        y="Velocity (in um/5 min)",
        title="Velocity of 10uM Treated CXT2 Cells")
print(plot)
```
